---
title: Data Visualization 
---

```{r}
library(tidyverse)

ggplot(msleep, aes(x= bodywt,y=awake, color= vore))+
  geom_point()+
  labs (
    x = "The body weight in kg",
    y = "The amount of time spent awake",
    title = "Mammals' body weight and their time spent awake"
  )

```


(Source from TidyTuesday; Updated sleep times and weights were taken from V. M. Savage and G. B. West. A quantitative, theoretical framework for understanding mammalian sleep. Proceedings of the National Academy of Sciences, 104 (3):1051-1056, 2007)

```{r}
ggplot(Orange, aes(x=age, y=circumference, color=Tree))+
  geom_point()+
  labs(
    x= "The age of the tree", 
    y="A numeric vector of trunk circumferences (mm)", 
    title= "The age of orange tree and their trunk circumferences"
  )
```

Source from TidyTuesday 
Draper, N. R. and Smith, H. (1998), Applied Regression Analysis (3rd ed), Wiley (exercise 24.N).

Pinheiro, J. C. and Bates, D. M. (2000) Mixed-effects Models in S and S-PLUS, Springer.



