---
title: Data Visualization 
---
```{r}

```

```{r}
library(tidyverse)
```

#Data Visualization on mammals' body weight and their time spent awake
```{r}
newdata<- msleep %>% filter(!is.na(vore)) |>
  filter(bodywt< 300)

ggplot(newdata, aes(x= bodywt, y=awake, color= vore))+
  geom_point(na.rm = TRUE)+
  labs (
    x = "The body weight in kg",
    y = "The amount of time spent awake",
    title = "Mammals' body weight and their time spent awake", 
    subtitle = "Excluding mammals that is hearvier than 300 kg/ outliners"
  )

```


(Source from TidyTuesday; from V. M. Savage and G. B. West.)






#Data Visualization on the age of orange tree and their trunk circumferences
```{r}
ggplot(Orange, aes(x=age, y=circumference, color=Tree))+
  geom_point()+
  labs(
    x= "The age of the tree", 
    y="A numeric vector of trunk circumferences (mm)", 
    title= "The age of orange tree and their trunk circumferences"
  )
```

(Source from Tidytuesday; Draper, N. R. and Smith, H. (1998), Applied Regression Analysis (3rd ed), Wiley (exercise 24.N).and Pinheiro, J. C. and Bates, D. M. (2000) Mixed-effects Models in S and S-PLUS, Springer.)



