---
title: "Presentation"
author: "Jessica Lin"
subtitle: "Dec 9"
format:
  revealjs:
    scrollable: true
    slide-number: true
    show-slide-number: all
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

## Project: Text Analysis

-   str\_\*() functions and regular expressions
-   Netflix Movies and TV Shows Data set
-   The data set comes from Kaggle w/ credit to Shivam Bansal
```{r, warning=FALSE, message=FALSE}
library("tidytuesdayR")
library("tidyverse")

netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-20/netflix_titles.csv')

show_col_types = FALSE
```

## Shivam Bansal

-   An experienced data science professional with a blend of full-stack engineering skills, pre-sales, and business skills
-   A director for smaller scale video contents
-   Data-sets for not just Netflix, but also many other video streaming platforms such as Amazon Prime, Disney+, and Hulu.

## Focus on US based productions
```{r}
library(dplyr)
library(stringr)

USfilms <- netflix_titles |> 
    filter(str_detect(country, "United States")) |> # Filter for US-based titles
  select(title, country, type) |> # Select only the relevant columns
  filter(!is.na(everything)) # Remove rows with NA values in any of the selected columns
```

## Organize the data to find first character 
```{r, warning=FALSE, message=FALSE}
N_firstword <- USfilms |>
filter(str_detect(title, "^[A-Z]")) |> # Filter titles starting with uppercase letters
  mutate(first_word = str_extract(title, "^\\w")) |> # Extract the first word (not just the first letter)
  group_by(first_word) |> # Group by the first word
  mutate(count = n()) # Add a count of occurrences within each group

print(N_firstword)
```

## ggplot 
```{r, warning=FALSE, message=FALSE}
library(ggplot2)

ggplot(N_firstword, aes(x= first_word, fill=type)) +
   geom_bar() +
  labs(title = "Occurrences of different first letter in US Netflix Productions",
       x = "First Letter of the title (A ~Z)",
       y = "Number of Occurrences") 
```

## Organize the data to find first digit 
```{r, warning=FALSE, message=FALSE}
N_firstdigit <- USfilms |>
  filter(str_detect(title, "^\\d+")) |>
  mutate(first_digit = str_extract(title, "^\\d")) |>
  group_by(first_digit) |>
  mutate(count = n())

print(N_firstdigit)
```

## ggplot
```{r, warning=FALSE, message=FALSE}
 ggplot(N_firstdigit, aes(x= first_digit, fill=type)) +
   geom_bar() +
  labs(title = "Occurrences of different first digit in US Netflix Productions",
       x = "First Digit (0~9)",
       y = "Number of Occurrences") 
```

## insight 
The analysis of first letters in US Netflix productions reveals some intriguing patterns:

-   Dominance of "T": The letter "T" is the most common starting letter for titles. This is likely due to the prevalence of titles that begin with "The," a common article in English.

-   Prevalence of the Digit "1": The digit "1" appears most frequently as the starting digit in both movies and TV shows.

-   Absence of Titles Starting with Certain Digits: There are no TV shows starting with the digit "2" or "4," possibly indicating a lack of popular content or marketing focus. Similarly, no movies begin with "7," which may reflect industry naming trends or a scarcity of suitable productions.

