---
title: "Project 4: SQL"
output:
  html_document:
    message: FALSE
    warning: FALSE
---

```{r}
library(dbplyr)
library(dplyr)
```

```{r}
library(RMariaDB)
con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")
```

```{sql connection=con_wai}
SHOW TABLES;
```

For this project, we are duplicate the Figure 1 from [Voss (2020)](https://pmc.ncbi.nlm.nih.gov/articles/PMC7093226/). We will query the WAI Database and build a data set in order to graph using ggplot.

```{sql connection=con_wai}
SELECT 
    Identifier, Instrument, 
    COUNT(DISTINCT CONCAT(SubjectNumber, Ear)) AS Unique_Ears,
    Frequency,
    AVG(Absorbance) AS MeanAbsorbance
FROM Measurements
WHERE Identifier IN ('Abur_2014', 'Feeney_2017', 'Groon_2015', 
                     'Lewis_2015', 'Liu_2008', 'Rosowski_2012', 
                     'Shahnaz_2006', 'Shaver_2013', 'Sun_2016', 
                     'Voss_1994', 'Voss_2010', 'Werner_2010')
GROUP BY Identifier, Instrument, Frequency
ORDER BY Frequency;
```

```{r}
graph_data <- "
SELECT 
    m.Identifier,
    m.Instrument,
    COUNT(DISTINCT CONCAT(m.SubjectNumber, m.Ear)) AS Unique_Ears, -- Count unique ears
    m.Frequency,
    AVG(m.Absorbance) AS MeanAbsorbance
FROM Measurements m
WHERE m.Identifier IN ('Abur_2014', 'Feeney_2017', 'Groon_2015', 
                        'Lewis_2015', 'Liu_2008', 'Rosowski_2012', 
                        'Shahnaz_2006', 'Shaver_2013', 'Sun_2016', 
                        'Voss_1994', 'Voss_2010', 'Werner_2010')
  AND m.Frequency > 200 -- Filter relevant frequencies
  AND m.Frequency < 8000 -- Align with the x-axis scale
GROUP BY m.Identifier, m.Instrument, m.Frequency
ORDER BY m.Identifier, m.Frequency;
"
```

```{r}
graph_plot <- dbGetQuery(con_wai, graph_data)

library(ggplot2)

ggplot(graph_plot, aes(x = Frequency, y = MeanAbsorbance, color = Identifier)) +
  geom_line() +
  scale_x_log10() + 
  labs(
    title = "Mean Absorbance from Each Publication in WAI Database",
    x = "Frequency (Hz)",
    y = "Mean Absorbance") +
    theme_minimal() 

```

The graph displays the mean absorbance (on y-axis) across different frequencies (Hz) (on x-axis) for multiple studies, each represented by a distinct line. The graph compares how absorbance patterns vary by frequency among studies using different systems.

# Part 2 - Specific Group of interest

For part 2, we are examining subjects with different race, especially Black and Caucasian.

```{sql connection=con_wai}
SELECT 
    Identifier,
    COUNT(*) AS Total_Subjects,
    COUNT(DISTINCT Race) AS Race_Groups,
    SUM(CASE WHEN Race = 'Unknown' THEN 1 ELSE 0 END) AS Unknown
FROM Subjects
GROUP BY Identifier
HAVING Race_Groups > 1 AND (Total_Subjects - Unknown) > 0; 
```

```{r}
query_race <- "
SELECT 
    Identifier,
    COUNT(*) AS Total_Subjects,
    COUNT(DISTINCT Race) AS Race_Groups,
    SUM(CASE WHEN Race = 'Unknown' THEN 1 ELSE 0 END) AS Unknown
FROM Subjects
GROUP BY Identifier
HAVING Race_Groups > 1 AND (Total_Subjects - Unknown) > 0;"

diverse_studies <- dbGetQuery(con_wai, query_race)

print(diverse_studies)

```

```{sql connection=con_wai}

SELECT 
    m.Identifier,       
    s.Race AS GroupName,  
    m.Frequency,         
    AVG(m.Absorbance) AS MeanAbsorbance 
FROM Measurements m
JOIN Subjects s 
    ON m.SubjectNumber = s.SubjectNumber
    AND m.Identifier = s.Identifier
WHERE m.Identifier IN ('Abur_2014', 'Feeney_2017', 'Groon_2015', 
                        'Lewis_2015', 'Liu_2008', 'Rosowski_2012', 
                        'Shahnaz_2006', 'Shaver_2013', 'Sun_2016', 
                        'Voss_1994', 'Voss_2010', 'Werner_2010') 
  AND m.Frequency > 200            
  AND m.Frequency < 8000
  AND m.Absorbance >= 0            
  AND s.Race IN ( 'Chinese', 'Caucasian')   
GROUP BY m.Identifier, s.Race, m.Frequency 
ORDER BY m.Identifier, s.Race, m.Frequency;

```

```{r}
query_race <- "

SELECT 
    m.Identifier,       
    s.Race AS GroupName,  
    m.Frequency,         
    AVG(m.Absorbance) AS MeanAbsorbance 
FROM Measurements m
JOIN Subjects s 
    ON m.SubjectNumber = s.SubjectNumber
    AND m.Identifier = s.Identifier
WHERE m.Identifier IN ('Abur_2014', 'Feeney_2017', 'Groon_2015', 
                        'Lewis_2015', 'Liu_2008', 'Rosowski_2012', 
                        'Shahnaz_2006', 'Shaver_2013', 'Sun_2016', 
                        'Voss_1994', 'Voss_2010', 'Werner_2010') 
  AND m.Frequency > 200            
  AND m.Frequency < 8000
  AND m.Absorbance >= 0            
  AND s.Race IN ( 'Black', 'Caucasian')  
GROUP BY m.Identifier, s.Race, m.Frequency 
ORDER BY m.Identifier, s.Race, m.Frequency;"


data_absorption <- dbGetQuery(con_wai, query_race)


head(data_absorption)
```

```{r}

ggplot(data_absorption, aes(x = Frequency, y = MeanAbsorbance, color = GroupName)) +
  geom_line() +
  facet_wrap(~ Identifier) +  
  scale_x_log10() +          
  labs(
    title = "Mean Absorbance by Ethnicity Across Multiple Studies",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    color = "Race"
  ) +
  theme_minimal()

```
